\chapter{Application cases}
\label{chap:use_cases}
	\par This chapter deals with every application case discussed in the study. As I explained three different area of interest mobile applications have been investigated: \textit{Weather}, \textit{Health \& Fitness} and \textit{Maps \& Navigation}. The application case sequence choice  has been made by following an increasing complexity order and an increasing amount of informations exchanged. \newline
	\par Generally the applications belonging to the first category do not require any form of user authentications, meaning that few private informations would be shared about the user. For this reason a single application has been taken under examination, that is \textit{iLMeteo}.\newline
	The application that falls in the second category, often have an higher degree of customization in terms of user profile, also implementing some features for user interaction in the application logic. This is the case of \textit{Pacer}, in which the investigation brought some really interesting results.\newline
	The last category is potentially the highest source of private user informations retrievable. Maps and Navigator applications require a continuous data exchange between client and server, in oder to deliver specific real time features, like the computation of the fastest path or the current traffic situation. Three different application have been taken in considerations, that are \textit{RadarBot}, \textit{Waze} and \textit{GoogleMaps}. \newline
	Obviously the higher the competitor is, the higher the analysis would be harder. This is the reason why there are so many concepts explained in the Fundamentals chapter (Section \ref{fundamentals}). Most of them are used not only to improve the application performance, but also to harden the security of the application itself. \newline
	\par The research study conducted application by application is reported in the following sections.
	
	\input{chapters/cases/iLMeteo.tex}
	
	\input{chapters/cases/Pacer.tex}
	
	\input{chapters/cases/RadarBot.tex}
	
	\input{chapters/cases/Waze.tex}
	
	\input{chapters/cases/GoogleMaps.tex}
	
	